### 2021-05-17 chatbot_classification/03.emnlp_2021/experiments.tex
modified:   experiments.tex

\subsubsection*{Are multi-task and joint training working?} 
First, we compare the four SFC variants shown in Table \ref{tbe:table2} to analyze the improvement
brought by multi-task and joint-model structure.

\begin{table}
  \begin{centering}
    \scalebox{0.65}{
      \begin{tabular}{|c|ccc|ccc|}
        \hline
        \multicolumn{1}{|c|}{\multirow{2}*{\textbf{Models}} }& \multicolumn{3}{c|}{\textbf{FRAMES}}& \multicolumn{3}{c|}{\textbf{ATIS}}\tabularnewline
        \cline{2-7}
        \multicolumn{1}{|c|}{} & $\textbf{10}$ & $\textbf{20}$ & $\textbf{50}$
        & $\textbf{10}$ & $\textbf{20}$ & $\textbf{50}$\tabularnewline
        \hline
        RoBERTa-large & \multirow{2}*{${0.3456}$} & \multirow{2}*{${0.4043}$} & \multirow{2}*{${0.4262}$} & \multirow{2}*{${0.9349}$} & \multirow{2}*{${0.9757}$} & \multirow{2}*{${0.9800}$}\tabularnewline
        (classification) & & & & & & \tabularnewline
        \hline
        RoBERTa-large & \multirow{3}*{${0.3520}$} & \multirow{3}*{${0.4088}$} & \multirow{3}*{${0.4353}$} & \multirow{3}*{${0.9560}$} & \multirow{3}*{${0.9706}$} & \multirow{3}*{$\textbf{0.9832}$}\tabularnewline
        (classification & & & & & & \tabularnewline
        w/ NER) & & & & & & \tabularnewline
        \hline
        Joint-SFC & \multirow{2}*{${0.3843}$} &
        \multirow{2}*{${0.4130}$} & \multirow{2}*{${0.4390}$} &
        \multirow{2}*{${0.9618}$} & \multirow{2}*{${0.9761}$} & 
        \multirow{2}*{${0.9825}$}\tabularnewline
        (w/o NER) & & & & & & \tabularnewline
        \hline
        Joint-SFC & \multirow{2}*{$\textbf{0.3925}$} &
        \multirow{2}*{$\textbf{0.4420}$} & \multirow{2}*{$\textbf{0.4456}$} &
        \multirow{2}*{$\textbf{0.9639}$} & \multirow{2}*{$\textbf{0.9784}$} & 
        \multirow{2}*{${0.9826}$}\tabularnewline
        (w/ NER) & & & & & & \tabularnewline
        \hline
      \end{tabular}
    }
  \end{centering}

@@

Comparing with training 2-stage SFC with multi-task, training with only task
1, namely the sentence pair similarity model, degrades by 0.8 percent
point on average, and training with only task 2, namely the top-K based
classification task, degrades by 0.45 percent point on average. These
degradation indicates multi-task training in 2-stage SFC is helpful for the
system performance. Besides, task 2 plays a relatively more important role in
multi-task learning, and this aligns with their optimal weight settings.

Joint-SFC consistently outperforms 2-stage SFC with multi-task training 
on 5 diverse datasets by 0.95 percents on average. 
This observation supports the idea that the joint model
structure can alleviate the limitation brought by the lack of interaction
between two stages. In following analysis, we will focus on the comparison
between joint-SFC and the other baseline models.

\subsubsection*{Is the fusion of classification and similarity models
helpful?} We compare joint-SFC with classification, similarity and NER joint model based baselines respectively.

@@

Regarding the classification models, joint-SFC
outperforms RoBERTa-large based, the strongest among all baselines, 
by 2.04 percent points on average F1 score over 5 datasets. 
Especially, ALBERT.xxlarge based is rather unstable in this scenario. Thus, 
we did multiple run with different settings and report the best ones here. 
Comparatively, RoBERTa based is the most stable and has the best performance 
n almost all experiment settings.

Regarding the similarity model, as analyzed above, we only choose the
RoBERTa-large based implementation as our baseline. Joint-SFC achieves more improvement of
7.09 percent points on average F1 score.

As for the intent classification and NER joint model, we choose the state-of-the-art RoBERTa-large model with NER loss as our strongest baseline. For ablation study, we did experiments on joint-SFC model with and without NER loss respectively. From Table \ref{tbe:addtional ner experiments}, we can see that our joint-SFC without NER loss still outperforms RoBERTa with NER loss baseline by 0.86 percent points on average. This indicates that the fusion of classification and similarity can even works better than fusing with additional NER information under few-shot chatbot scenario. Moreover, after adding NER loss to joint-SFC, it achieves further improvement by 1.66 percent points over RoBERTa with NER loss baseline. Especially for FRAMES dataset, which is a much more difficult task in comparison with ATIS, our joint-SFC with NER loss achieves a bigger improvement of 2.8 percent points on average. In this way, we show the compatibility of our joint-SFC model with extra semantic information, as well as the stable outstanding performance of joint-SFC when facing hard tasks with few training data.

@@

\subsubsection*{Does joint-SFC improve the classification in stage 1?}
We compare the top-1 prediction and top-5 candidate labels in stage 1 from
2-stage SFC and joint-SFC in Table \ref{tbe:top1_5_accuracy}. We should note
that 2-stage SFC does not optimize the stage 1 model, which is a RoBERTa
based classification model.

The joint training does not improve the classification performance in stage 1,
measured by top-1 accuracy. However, it improves the quality of top-5
candidate class labels, which can further improve the effect of sentence pair
similarity training. The reason is that, a classification model
inherently optimizes the objective loss, 0-1 error of top-1 here; the sentence
pair similarity model in the joint-SFC poses more positive effect in optimizing the 
candidate labels.

\subsubsection*{How does training sample size influence SFC?} 
We analyze the overall improvement trend of joint-SFC under various few-shot
settings, and show results in Fig. \ref{fig:trend}.

Our main focus is to study chatbot building in common situation during real-world
application where only a few sample sentences are available for each class.
The experimental results for the 3 diverse intent detection datasets
(ClINC150, BANKING77, HWU64) under few-shot settings, 5 to 50 samples per
class, indicate that our proposed joint-SFC can achieve 2.47 percents average
improvement over one of the most powerful baseline models, which is
RoBERTa-large classification model. Moreover, the improvement in F1 score
becomes more and more prominent with the decrease of data sample number per
class.

\begin{figure}[t]
  \begin{centering}
    \includegraphics[scale=0.2]{picture/improvement_trend.jpg} 
    \par
  \end{centering}
  \caption{
    Improvements from Joint SFC over RoBERTa based classification with
    different size of training data.
  }
  \label{fig:trend}
\end{figure}

Especially, in the most extreme few-shot setting with only 5 training samples
per class our joint-SFC achieves 4.97 percent points improvements on average
over RoBERTa-large classification model. This observation indicates that SFC
is especially well-performed when being applied during the initial stage of
building task-specific chatbot system, where the amount of data sample for
each class label is extremely scarce.

Furthermore, our proposed joint SFC can also achieve 2.04 percents average
improvement over RoBERTa-large classification model on all the data settings
including the ones with large scale, 30 to 50 samples per class. This
indicates that joint-SFC can still steadily outperforms pretrained model based
baseline even if the data size grows bigger. However, the improvement is still
more prominent when the data is scarce, which makes SFC an excellent model for
low-resource chatbot scenario.

\subsubsection*{How does joint-SFC relate to the label embedding based LEAM?}
LEAM performs consistently better than TextCNN by a large margin in all
datasets. TextCNN is practically efficiently in task-specific chat
applications, and LEAM shows its power in modeling labeling information.

LEAM does not perform as well as SFCs, since the first intuitive reason is it
does not use pretrained model as encoding module; and another critical reason
is in task-specific chat applications, it is common to have many intents
that are close to some others with a minor difference, and it is hard and even
impossible to name each intent with a short clear name. One candidate
solution is setting a most standard sample as the label of that class. When
using non-pretrained models base LEAM, this is applicable. Yet when using
pretrained model based, as the all labels can be hundreds of thousands, then
this can not be accommodated in a poplar 32 G Tesla V100 GPU. Actually, joint
SFC can be kind of understood as a generalization form of LEAM. In the
scenario when there is no clear class label, the relationship between a sample
and a class label is implicitly encoded as that between a sample and another
sample from the same class, and this turns into a sentence pair similarity
model.

>>>>

\subsubsection*{Multi-task and joint training} 
\begin{table}
  \begin{centering}
    \scalebox{0.65}{
      \begin{tabular}{|c|ccc|ccc|}
        \hline
        \multicolumn{1}{|c|}{\multirow{2}*{\textbf{Models}} }& \multicolumn{3}{c|}{\textbf{FRAMES}}& \multicolumn{3}{c|}{\textbf{ATIS}}\tabularnewline
        \cline{2-7}
        \multicolumn{1}{|c|}{} & $\textbf{10}$ & $\textbf{20}$ & $\textbf{50}$
        & $\textbf{10}$ & $\textbf{20}$ & $\textbf{50}$\tabularnewline
        \hline
        RoBERTa-large & \multirow{2}*{${0.3456}$} & \multirow{2}*{${0.4043}$} & \multirow{2}*{${0.4262}$} & \multirow{2}*{${0.9349}$} & \multirow{2}*{${0.9757}$} & \multirow{2}*{${0.9800}$}\tabularnewline
        (classification) & & & & & & \tabularnewline
        \hline
        RoBERTa-large & \multirow{3}*{${0.3520}$} & \multirow{3}*{${0.4088}$} & \multirow{3}*{${0.4353}$} & \multirow{3}*{${0.9560}$} & \multirow{3}*{${0.9706}$} & \multirow{3}*{$\textbf{0.9832}$}\tabularnewline
        (classification & & & & & & \tabularnewline
        w/ NER) & & & & & & \tabularnewline
        \hline
        Joint-SFC & \multirow{2}*{${0.3843}$} &
        \multirow{2}*{${0.4130}$} & \multirow{2}*{${0.4390}$} &
        \multirow{2}*{${0.9618}$} & \multirow{2}*{${0.9761}$} & 
        \multirow{2}*{${0.9825}$}\tabularnewline
        (w/o NER) & & & & & & \tabularnewline
        \hline
        Joint-SFC & \multirow{2}*{$\textbf{0.3925}$} &
        \multirow{2}*{$\textbf{0.4420}$} & \multirow{2}*{$\textbf{0.4456}$} &
        \multirow{2}*{$\textbf{0.9639}$} & \multirow{2}*{$\textbf{0.9784}$} & 
        \multirow{2}*{${0.9826}$}\tabularnewline
        (w/ NER) & & & & & & \tabularnewline
        \hline
      \end{tabular}
    }
  \end{centering}

@@

Comparing with training 2-stage SFC with multi-task Table \ref{tbe:table2}, training with only task 1, namely the sentence pair similarity model, degrades by 0.8 percentage point on average, and training with only task 2, namely the top-K based classification task, degrades by 0.45 percentage point on average. 
These degradation indicates multi-task training in 2-stage SFC is helpful for the system performance. Besides, task 2 plays a relatively more important role in multi-task learning, and this aligns with their optimal weight settings.

Joint-SFC consistently outperforms 2-stage SFC with multi-task training on 5 diverse datasets by 0.95 percentage point on average. 
Though a joint model structure may bring extra complexity to the system, it does alleviate the error propagation from the first stage and improves the final performance.

In following analysis, we will focus on the comparison between joint-SFC and the other baseline models.

\subsubsection*{Fusion of classification and similarity models} 
%We compare joint-SFC with classification, similarity and NER joint model based baselines respectively.

@@

Regarding the classification models, joint-SFC outperforms RoBERTa-large based, the strongest among all baselines, by 2.04 percentage points on average F1 score over 5 datasets. 
Especially, ALBERT.xxlarge based is rather unstable in this short sentence task. 
Thus, we run multiple times with different settings and report the best ones here. 
Comparatively, RoBERTa based is the most stable and has the best performance almost all experiment settings.

Regarding the similarity model, as analyzed above, we only choose the RoBERTa-large based implementation as our baseline. 
Joint-SFC achieves more improvement of 7.09 percentage points on average F1 score. 
This is also understandable, since RoBERTa-large based similarity model does not and also can not train towards the task specific goal.

@@

\subsubsection*{Joint-SFC improves stage 1?}
We explore the quality of predicted candidate labels in stage 1 from 2-stage SFC and joint-SFC respectively. 
We should note that 2-stage SFC does not optimize the stage 1 model, namely a RoBERTa based classification model.

From Table \ref{tbe:top1_5_accuracy}, the joint training does not improve the top-1 classification accuracy in stage 1, yet it improves the top-5 accuracy.
The reason is that, a classification model inherently optimizes the objective loss, 0-1 error of top-1 here; the sentence pair similarity model in joint-SFC poses more positive effect in optimizing the candidate labels.

\subsubsection*{Supportive of extra NER loss}
Many previous work shows joint training with a NER loss improves the intent classification task as well, observed again in our baselines and SFCs in Table \ref{tbe:addtional ner experiments}.
We choose the state-of-the-art RoBERTa-large model with NER loss as a strong baseline.

We can see that our joint-SFC without NER loss still outperforms RoBERTa with NER loss baseline by 0.86 percentage points on average. 
This indicates that the fusion of classification and similarity models can even works better than fusing with additional NER information under few-shot chatbot scenario.

Moreover, the adding NER loss to joint-SFC further achieves improvement of 1.66 percentage points over RoBERTa with NER loss baseline. 
Especially on FRAMES dataset, which is a much more difficult task in comparison with ATIS based on performance of all systems on it, our joint-SFC with NER loss achieves a bigger improvement of 2.8 percentage points on average.

\subsubsection*{Different training sample size} 
We analyze the overall improvement trend of joint-SFC under various few-shot settings, and show results in Fig. \ref{fig:trend}.

Our main focus is on studying chatbot construction in real world applications, where each class has only a small number of example sentences.
The experimental results for the 3 diverse intent detection datasets (ClINC150, BANKING77, HWU64) under few-shot settings, 5 to 50 samples per class, indicate that our proposed joint-SFC can achieve average 2.47 percentage points improvement over one of the most powerful baseline models, which is RoBERTa-large classification model. 
Moreover, the improvement in F1 score becomes more and more prominent with the decrease of data sample number per class.

\begin{figure}[t]
  \begin{centering}
    \includegraphics[scale=0.2]{picture/improvement_trend.jpg} 
    \par
  \end{centering}
  \caption{
    Improvements from joint SFC over RoBERTa based classification with different training size.
  }
  \label{fig:trend}
\end{figure}

Especially, in the most extreme setting with only 5 training samples per class, joint-SFC achieves 4.97 percentage points improvements on average over RoBERTa-large classification model. 
Joint SFC is a better choice applied into the early stage of building a task-specific chatbot system, since label data is extremely scarce.

Furthermore, joint SFC also achieves average 2.04 percentage points improvement over RoBERTa-large classification model on all the data settings. 
However, the improvement is still more prominent when the data is scarce, which makes SFC an excellent model for low-resource chatbot scenario.

\subsubsection*{Relation to the label embedding based LEAM}
LEAM shows its superiority over TextCNN in all datasets in our experiments, by modeling labeling information.

LEAM does not perform as well as SFCs, since the first intuitive reason is it does not use a pretrained model as its input encoding module; 
another critical reason is in task-specific chat applications, it is common to have many similar intents, and it becomes hard and even impossible to name each intent with a short clear name to feed into LEAM model.
One candidate solution is setting a most standard sample as the label of the class. 
When using non-pretrained models base LEAM, this is applicable. 
Yet when using a pretrained model based, as all labels can be hundreds of thousands, then the training can not be accommodated in a poplar 32 G Tesla V100 GPU.

Actually, joint SFC can be kind of understood as a generalization form of LEAM. 
When there is no clear class labels, the relationship between a sample and a class label is implicitly encoded as that between a sample and another sample from the same class, and this turns into a sentence pair similarity model.
